plyranges: a grammar for transforming genomics data

### 2018-04-20 paper.Rmd
add acknowledgments

     | Fluent                   | IRanges       | bedtools |
     |--------------------------+---------------+----------|
     | stretch(anchor_start())  | -             | slop -r  |
     | stretch(anchor_end())    | -             | slop -l  |
     | stretch(anchor_center()) | +()           | slop -b  |
     | set_width(anchor_*)      | resize(fix=*) | -        |
     | shift_right()            | shift()       | shift -s |
     | shift_left()             | -             | -        |
     | shift_up/downstream()    | -             | -        |

@@

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tibble)
feature_tbl <- tribble(
  ~Verb, ~Action, ~Description,
  "`mutate()`", "modifies columns", "Takes a GRanges object and a set of name-value pairs and generates a new GRanges object that with modified or new metadata columns or modified core components (start, end, width, seqnames, strand).",
 "`filter()`", "subset rows", "Takes a GRanges object and a set of logical expressions and restricts the GRanges object to where all logical expressions are true.",
 "`summarise()`", "aggregate one or more columns by a function", "Takes a GRanges object and a set of name-value pairs and aggregates the GRanges according to functions evaluated in the name-value pairs.",
"`group_by()`", "partition a GRanges object by a column", "Create an implicit grouping of GRanges object by one or more columns. This modifies the actions of verbs, so they are performed on each partition created by the grouping.",
"`select()`", "select columns of a GRanges objects", "Drop or reorder metadata columns.",
"`arrange()`", "sort a GRanges object by columns", "Sort a GRanges object by named variables",
"`disjoin_ranges()`", "aggregate one more columns of a GRanges object over the union of end coordinates", "Takes a GRanges object and a set of name-value pairs and expands the GRanges object by taking the union of end coordinates. For each row that was included to form the final union, the aggregation function defined by a name-value pair is called.",
"`reduce_ranges()`", "aggregate one more columns of a GRanges object by merging overlapping and neighbouring ranges.", "Takes a GRanges object and a set of name-value pairs and reduces the GRanges object by merging overlapping and neighouring ranges. For each row that was included to form merged range, the aggregation function defined by a name-value pair is called."
)

knitr::kable(feature_tbl, booktabs = TRUE)
```

@@

```{r}
h1_bam <- read_bam("h1.bam") %>%
  filter(mapq > 20) 
```

@@

```{r}
bins <- tile(h1_bam, width = 10000)
```

@@

```{r}
h1_bam_read_summary <- h1_bam %>%
  group_by_overlaps(bins) %>%
  summarise(n_reads = n())
```

@@

>>>>

@@

```{r, echo = FALSE, eval = FALSE}
library(tibble)
feature_tbl <- tribble(
  ~Verb, ~Action, ~Description,
  "`mutate()`", "modifies columns", "Takes a GRanges object and a set of name-value pairs and generates a new GRanges object that with modified or new metadata columns or modified core components (start, end, width, seqnames, strand).",
 "`filter()`", "subset rows", "Takes a GRanges object and a set of logical expressions and restricts the GRanges object to where all logical expressions are true.",
 "`summarise()`", "aggregate one or more columns by a function", "Takes a GRanges object and a set of name-value pairs and aggregates the GRanges according to functions evaluated in the name-value pairs.",
"`group_by()`", "partition a GRanges object by a column", "Create an implicit grouping of GRanges object by one or more columns. This modifies the actions of verbs, so they are performed on each partition created by the grouping.",
"`select()`", "select columns of a GRanges objects", "Drop or reorder metadata columns.",
"`arrange()`", "sort a GRanges object by columns", "Sort a GRanges object by named variables",
"`disjoin_ranges()`", "aggregate one more columns of a GRanges object over the union of end coordinates", "Takes a GRanges object and a set of name-value pairs and expands the GRanges object by taking the union of end coordinates. For each row that was included to form the final union, the aggregation function defined by a name-value pair is called.",
"`reduce_ranges()`", "aggregate one more columns of a GRanges object by merging overlapping and neighbouring ranges.", "Takes a GRanges object and a set of name-value pairs and reduces the GRanges object by merging overlapping and neighouring ranges. For each row that was included to form merged range, the aggregation function defined by a name-value pair is called."
)

knitr::kable(feature_tbl, format = "latex")
```

@@

```{r, eval = FALSE}
h1_bam <- read_bam("h1.bam") %>%
  filter(mapq > 20) 
```

@@

```{r, eval = FALSE}
bins <- tile(h1_bam, width = 10000)
```

@@

```{r, eval = FALSE}
h1_bam_read_summary <- h1_bam %>%
  group_by_overlaps(bins) %>%
  summarise(n_reads = n())
```

@@

# Acknowledgements

We would like to thank Dr Matthew Ritchie at the Walter and Eliza Hall 
Institute and Dr Paul Harrison for their feedback on earlier drafts of
this work. We would also like to thank Lori Shepherd and Hèrve Pages
for the code review they performed. This report was written with
`knitr` @R-knitr and the figures were made with `ggbio` @R-ggbio.
